\chapter{How to Get to Germany / 如何前往德国 / Wie man nach Deutschland kommt}

\section{Introduction / 介绍 / Einführung}

This chapter provides comprehensive information about visa types, immigration procedures, and requirements for Chinese entrepreneurs planning to establish a business in Germany.

本章为计划在德国创业的中国企业家提供有关签证类型、移民程序和要求的全面信息。

Dieses Kapitel bietet umfassende Informationen zu Visa-Typen, Einwanderungsverfahren und Anforderungen für chinesische Unternehmer, die planen, in Deutschland ein Unternehmen zu gründen.

\section{Visa Types for Entrepreneurs / 企业家签证类型 / Visa-Typen für Unternehmer}

\subsection{Self-Employment Visa (Freelancer Visa) / 自雇签证 / Selbständigenvisum}

The self-employment visa (Aufenthaltserlaubnis zur selbständigen Tätigkeit) is designed for individuals who wish to establish and operate their own business in Germany. This visa category has become more accessible following the 2020 Skilled Immigration Law (Fachkräfteeinwanderungsgesetz), which simplified entry requirements for entrepreneurs.

自雇签证（Aufenthaltserlaubnis zur selbständigen Tätigkeit）专为希望在德国建立和经营自己企业的个人设计。随着2020年《技术移民法》（Fachkräfteeinwanderungsgesetz）的出台，该签证类别变得更加容易获得，该法律简化了企业家的入境要求。

Das Selbständigenvisum (Aufenthaltserlaubnis zur selbständigen Tätigkeit) ist für Personen gedacht, die in Deutschland ein eigenes Unternehmen gründen und betreiben möchten. Diese Visumkategorie wurde durch das Fachkräfteeinwanderungsgesetz von 2020 zugänglicher, das die Einreiseanforderungen für Unternehmer vereinfachte.

\textbf{Key Requirements:}
\begin{itemize}
    \item \textbf{Business Plan}: A comprehensive business plan written in German or professionally translated is mandatory. The plan must demonstrate economic interest, viability, and sufficient funding. It should include market analysis, financial projections for 3 years, and proof that the business will create jobs or contribute to innovation.
    \item 商业计划：必须提供用德语撰写或专业翻译的综合商业计划。该计划必须证明经济利益、可行性和充足的资金。应包括市场分析、3年财务预测，以及证明该企业将创造就业机会或促进创新的证据。
    \item Businessplan: Ein umfassender Businessplan auf Deutsch oder professionell übersetzt ist obligatorisch. Der Plan muss wirtschaftliches Interesse, Tragfähigkeit und ausreichende Finanzierung nachweisen.
    
    \item \textbf{Financial Requirements}: You must prove sufficient capital to cover startup costs and living expenses. While there's no fixed minimum, authorities typically expect at least €25,000-50,000 for initial capital, plus proof of ability to support yourself and family for the first year (approximately €12,000-15,000 per person annually).
    \item 财务要求：您必须证明有足够的资金来支付启动成本和生活费用。虽然没有固定的最低限额，但当局通常期望至少有25,000-50,000欧元的初始资本，以及证明有能力在第一年支持自己和家人的证据（每人每年约12,000-15,000欧元）。
    \item Finanzielle Anforderungen: Sie müssen ausreichend Kapital nachweisen, um Gründungskosten und Lebenshaltungskosten zu decken. Typischerweise erwarten die Behörden mindestens 25.000-50.000 Euro Startkapital.
    
    \item \textbf{Economic Interest Assessment}: The local Chamber of Commerce (IHK) or economic development agency evaluates whether your business serves the economic interests of Germany. Factors include job creation potential, innovation contribution, and market demand.
    \item 经济利益评估：当地商会（IHK）或经济发展机构评估您的企业是否服务于德国的经济利益。因素包括创造就业的潜力、创新贡献和市场需求。
    \item Wirtschaftliches Interesse: Die örtliche IHK oder Wirtschaftsförderung bewertet, ob Ihr Unternehmen den wirtschaftlichen Interessen Deutschlands dient.
    
    \item \textbf{Professional Qualifications}: While not always mandatory, relevant education, work experience, or professional certifications strengthen your application.
    \item 专业资格：虽然不总是强制性的，但相关的教育、工作经验或专业认证会加强您的申请。
    \item Berufsqualifikationen: Relevante Bildung, Berufserfahrung oder Zertifikate stärken Ihre Bewerbung.
\end{itemize}

\textbf{Application Process:}
\begin{enumerate}
    \item Submit application at German embassy/consulate in China with all required documents
    \item 在中国德国大使馆/领事馆提交申请及所有所需文件
    \item Antrag bei der deutschen Botschaft/Konsulat in China einreichen
    \item Initial review typically takes 4-8 weeks
    \item 初步审查通常需要4-8周
    \item Erste Prüfung dauert typischerweise 4-8 Wochen
    \item If approved, receive entry visa (valid 3-6 months)
    \item 如果获得批准，将获得入境签证（有效期3-6个月）
    \item Bei Genehmigung erhält man ein Einreisevisum (3-6 Monate gültig)
    \item Enter Germany and apply for residence permit at local Foreigners' Office (Ausländerbehörde) within 90 days
    \item 进入德国并在90天内在当地外国人办公室（Ausländerbehörde）申请居留许可
    \item Nach Einreise innerhalb von 90 Tagen bei der Ausländerbehörde Aufenthaltserlaubnis beantragen
\end{enumerate}

\subsection{Business Investment Visa / 商业投资签证 / Geschäftsinvestitionsvisum}

The business investment visa is suitable for entrepreneurs making substantial capital investments in Germany. This pathway is particularly relevant for those establishing GmbH (Limited Liability Company) or larger corporate structures.

商业投资签证适用于在德国进行大量资本投资的企业家。此途径特别适用于建立GmbH（有限责任公司）或更大的公司结构。

Das Geschäftsinvestitionsvisum eignet sich für Unternehmer, die erhebliche Kapitalinvestitionen in Deutschland tätigen. Dieser Weg ist besonders relevant für die Gründung von GmbH oder größeren Unternehmensstrukturen.

\textbf{Investment Requirements:}
\begin{itemize}
    \item \textbf{Minimum Capital}: For GmbH formation, minimum share capital of €25,000 is required, with at least €12,500 paid-in at registration. However, authorities typically expect additional working capital (€50,000-100,000+) to demonstrate serious commitment and operational viability.
    \item 最低资本：对于GmbH的成立，需要最低股本25,000欧元，其中至少12,500欧元在注册时缴付。然而，当局通常期望额外的营运资金（50,000-100,000欧元以上）以证明认真的承诺和运营可行性。
    \item Mindestkapital: Für GmbH-Gründung sind mindestens 25.000 Euro Stammkapital erforderlich, davon mindestens 12.500 Euro bei Registrierung eingezahlt.
    
    \item \textbf{Job Creation}: While not always strictly required, demonstrating potential to create jobs for German or EU citizens significantly strengthens the application. A business plan showing 2-5 jobs within 2-3 years is typically viewed favorably.
    \item 创造就业：虽然不总是严格要求，但证明为德国或欧盟公民创造就业的潜力会显著加强申请。显示在2-3年内创造2-5个工作岗位的商业计划通常会被看好。
    \item Arbeitsplatzschaffung: Der Nachweis des Potenzials zur Schaffung von Arbeitsplätzen für deutsche oder EU-Bürger stärkt die Bewerbung erheblich.
    
    \item \textbf{Economic Interest}: The investment must serve Germany's economic interests. Factors evaluated include: innovation contribution, regional economic development, export potential, technology transfer, and filling market gaps.
    \item 经济利益：投资必须服务于德国的经济利益。评估的因素包括：创新贡献、区域经济发展、出口潜力、技术转让和填补市场空白。
    \item Wirtschaftliches Interesse: Die Investition muss den wirtschaftlichen Interessen Deutschlands dienen.
\end{itemize}

\textbf{Application Advantages:}
\begin{itemize}
    \item Faster processing for substantial investments (€100,000+)
    \item 大额投资（10万欧元以上）的更快处理
    \item Schnellere Bearbeitung bei erheblichen Investitionen (100.000+ Euro)
    \item Potential for permanent residence after 3 years of successful operation
    \item 成功运营3年后可能获得永久居留
    \item Möglichkeit der Niederlassungserlaubnis nach 3 Jahren erfolgreichem Betrieb
    \item Family members can accompany on dependent visas
    \item 家庭成员可以陪同，持家属签证
    \item Familienangehörige können mit abhängigen Visa begleiten
\end{itemize}

\subsection{Blue Card EU / 欧盟蓝卡 / Blaue Karte EU}

The EU Blue Card is designed for highly qualified professionals with university degrees. While primarily for employees, it can be relevant for entrepreneurs who also work in technical or managerial roles within their companies.

欧盟蓝卡专为拥有大学学位的高素质专业人士设计。虽然主要面向雇员，但对于在其公司内担任技术或管理职务的企业家也可能相关。

Die Blaue Karte EU ist für hochqualifizierte Fachkräfte mit Hochschulabschluss gedacht. Während sie hauptsächlich für Arbeitnehmer gedacht ist, kann sie für Unternehmer relevant sein, die auch technische oder Managementrollen in ihren Unternehmen ausüben.

\textbf{Requirements:}
\begin{itemize}
    \item \textbf{Education}: Recognized university degree (minimum 3-year program) or equivalent professional qualification
    \item 教育：认可的大学学位（至少3年制）或同等专业资格
    \item Bildung: Anerkannter Hochschulabschluss (mindestens 3 Jahre) oder gleichwertige Berufsqualifikation
    
    \item \textbf{Salary Threshold}: Annual gross salary of at least €58,400 (2024) for regular professions, or €45,552 for shortage occupations (STEM fields, IT, healthcare, engineering)
    \item 薪资门槛：普通职业年总收入至少58,400欧元（2024年），或紧缺职业（STEM领域、IT、医疗、工程）45,552欧元
    \item Gehaltsschwelle: Jahresbruttogehalt von mindestens 58.400 Euro (2024) für reguläre Berufe, oder 45.552 Euro für Mangelberufe
    
    \item \textbf{Job Offer}: Valid employment contract or binding job offer from German employer (can be your own company if properly structured)
    \item 工作机会：来自德国雇主的有效雇佣合同或具有约束力的工作机会（如果结构合理，可以是您自己的公司）
    \item Arbeitsangebot: Gültiger Arbeitsvertrag oder bindendes Arbeitsangebot von deutschem Arbeitgeber
\end{itemize}

\textbf{Advantages:}
\begin{itemize}
    \item Fast-track to permanent residence: After 33 months (or 21 months with B1 German level)
    \item 快速获得永久居留：33个月后（或B1德语水平21个月后）
    \item Schneller Weg zur Niederlassungserlaubnis: Nach 33 Monaten (oder 21 Monaten mit B1-Deutsch)
    \item Family members receive work permits automatically
    \item 家庭成员自动获得工作许可
    \item Familienangehörige erhalten automatisch Arbeitserlaubnisse
    \item Can change employers after 18 months without new approval
    \item 18个月后可以更换雇主，无需新批准
    \item Nach 18 Monaten Arbeitgeberwechsel ohne neue Genehmigung möglich
\end{itemize}

\section{Application Process / 申请流程 / Antragsverfahren}

\subsection{Visa Application Process Flow / 签证申请流程 / Visumantragsprozess-Flussdiagramm}

Figure~\ref{fig:visa_process} illustrates the complete visa application process from initial preparation to obtaining a residence permit in Germany.

图~\ref{fig:visa_process}说明了从初始准备到在德国获得居留许可的完整签证申请流程。

Abbildung~\ref{fig:visa_process} veranschaulicht den vollständigen Visumantragsprozess von der ersten Vorbereitung bis zur Erteilung einer Aufenthaltserlaubnis in Deutschland.

\begin{figure}[ht]
\centering
\begin{tikzpicture}[
    node distance=1.5cm,
    auto,
    decision/.style={diamond, draw, fill=blue!20, text width=3cm, text centered, aspect=2},
    process/.style={rectangle, draw, fill=green!20, text width=3.5cm, text centered, rounded corners},
    startend/.style={ellipse, draw, fill=red!20, text width=2.5cm, text centered},
    arrow/.style={->, >=stealth, thick}
]
    % Start
    \node [startend] (start) {Start / 开始 / Start};
    
    % Preparation phase
    \node [process, below of=start] (prep) {Prepare Documents / 准备文件 / Dokumente vorbereiten};
    \node [process, below of=prep] (plan) {Develop Business Plan / 制定商业计划 / Businessplan erstellen};
    \node [process, below of=plan] (funds) {Proof of Funds / 资金证明 / Mittel nachweisen};
    
    % Application
    \node [process, below of=funds] (apply) {Submit Application / 提交申请 / Antrag einreichen};
    \node [process, below of=apply] (review) {Embassy Review / 大使馆审查 / Botschaftsprüfung};
    
    % Decision
    \node [decision, below of=review] (decision) {Approved? / 批准? / Genehmigt?};
    
    % Approved path
    \node [process, right of=decision, xshift=3cm] (visa) {Receive Entry Visa / 获得入境签证 / Einreisevisum erhalten};
    \node [process, below of=visa] (enter) {Enter Germany / 进入德国 / Nach Deutschland einreisen};
    \node [process, below of=enter] (register) {Register at Bürgeramt / 在Bürgeramt登记 / Bei Bürgeramt anmelden};
    \node [process, below of=register] (permit) {Apply Residence Permit / 申请居留许可 / Aufenthaltserlaubnis beantragen};
    \node [startend, below of=permit] (success) {Residence Permit / 居留许可 / Aufenthaltserlaubnis};
    
    % Rejected path
    \node [process, left of=decision, xshift=-3cm] (reject) {Application Rejected / 申请被拒 / Antrag abgelehnt};
    \node [startend, below of=reject] (end) {Appeal/Reapply / 上诉/重新申请 / Widerspruch/Neu beantragen};
    
    % Arrows
    \draw [arrow] (start) -- (prep);
    \draw [arrow] (prep) -- (plan);
    \draw [arrow] (plan) -- (funds);
    \draw [arrow] (funds) -- (apply);
    \draw [arrow] (apply) -- (review);
    \draw [arrow] (review) -- (decision);
    \draw [arrow] (decision) -- node[above] {Yes / 是 / Ja} (visa);
    \draw [arrow] (decision) -- node[above] {No / 否 / Nein} (reject);
    \draw [arrow] (visa) -- (enter);
    \draw [arrow] (enter) -- (register);
    \draw [arrow] (register) -- (permit);
    \draw [arrow] (permit) -- (success);
    \draw [arrow] (reject) -- (end);
\end{tikzpicture}
\caption{Visa Application Process Flow / 签证申请流程图 / Visumantragsprozess-Flussdiagramm}
\label{fig:visa_process}
\end{figure}

\subsection{Step-by-Step Guide / 分步指南 / Schritt-für-Schritt-Anleitung}

The visa application process typically takes 2-4 months from initial submission to approval. Here's a detailed timeline and process:

签证申请过程通常从初次提交到批准需要2-4个月。以下是详细的时间表和流程：

Der Visumantragsprozess dauert typischerweise 2-4 Monate von der ersten Einreichung bis zur Genehmigung. Hier ist ein detaillierter Zeitplan und Prozess:

\begin{enumerate}
    \item \textbf{Preparation Phase (4-8 weeks before application)} / \textbf{准备阶段（申请前4-8周）} / \textbf{Vorbereitungsphase (4-8 Wochen vor Antrag)}
    \begin{itemize}
        \item Gather all personal documents (passport, birth certificate, educational certificates, marriage certificate if applicable)
        \item 收集所有个人文件（护照、出生证明、教育证书、如适用结婚证）
        \item Alle persönlichen Dokumente sammeln (Reisepass, Geburtsurkunde, Bildungszertifikate, Heiratsurkunde falls zutreffend)
        \item Develop comprehensive business plan in German (or have professionally translated)
        \item 用德语制定综合商业计划（或进行专业翻译）
        \item Umfassenden Businessplan auf Deutsch erstellen (oder professionell übersetzen lassen)
        \item Obtain proof of funds (bank statements showing sufficient capital for 12+ months)
        \item 获得资金证明（显示足够12个月以上资金的银行对账单）
        \item Nachweis der Mittel beschaffen (Kontoauszüge mit ausreichendem Kapital für 12+ Monate)
        \item Secure health insurance coverage valid in Germany
        \item 获得在德国有效的健康保险
        \item Krankenversicherungsschutz für Deutschland sichern
        \item Research and contact local Chamber of Commerce (IHK) for preliminary assessment
        \item 研究并联系当地商会（IHK）进行初步评估
        \item Lokale IHK für Vorabprüfung recherchieren und kontaktieren
    \end{itemize}
    
    \item \textbf{Business Plan Development (Critical Component)} / \textbf{商业计划制定（关键组成部分）} / \textbf{Businessplan-Entwicklung (Kritische Komponente)}
    
    Your business plan must be comprehensive and written in German. Key sections required:
    
    您的商业计划必须全面且用德语撰写。需要的关键部分：
    
    Ihr Businessplan muss umfassend und auf Deutsch verfasst sein. Erforderliche Schlüsselabschnitte:
    \begin{itemize}
        \item Executive Summary (1 page maximum): Business concept, unique value proposition, target market
        \item 执行摘要（最多1页）：商业概念、独特价值主张、目标市场
        \item Executive Summary (max. 1 Seite): Geschäftskonzept, einzigartiges Wertversprechen, Zielmarkt
        \item Personal Background: Education, work experience, relevant skills, motivation for starting business in Germany
        \item 个人背景：教育、工作经验、相关技能、在德国创业的动机
        \item Persönlicher Hintergrund: Bildung, Berufserfahrung, relevante Fähigkeiten, Motivation
        \item Product/Service Description: Detailed description with competitive differentiation
        \item 产品/服务描述：详细描述，包括竞争差异化
        \item Produkt-/Dienstleistungsbeschreibung: Detaillierte Beschreibung mit Wettbewerbsdifferenzierung
        \item Market Analysis: Target customers, market size, competition analysis, market entry strategy
        \item 市场分析：目标客户、市场规模、竞争分析、市场进入策略
        \item Marktanalyse: Zielkunden, Marktgröße, Wettbewerbsanalyse, Markteintrittsstrategie
        \item Financial Projections: 3-year revenue/expense forecasts, break-even analysis, capital requirements
        \item 财务预测：3年收入/支出预测、盈亏平衡分析、资本要求
        \item Finanzprognosen: 3-Jahres-Umsatz-/Ausgabenprognosen, Break-Even-Analyse, Kapitalanforderungen
        \item Marketing Strategy: Customer acquisition channels, pricing strategy, promotion plan
        \item 营销策略：客户获取渠道、定价策略、推广计划
        \item Marketingstrategie: Kundenakquisitionskanäle, Preisstrategie, Werbeplan
    \end{itemize}
    
    \item \textbf{Application Submission} / \textbf{提交申请} / \textbf{Antragseinreichung}
    \begin{itemize}
        \item Schedule appointment at German embassy/consulate in China (Beijing, Shanghai, Guangzhou, Chengdu)
        \item 在中国德国大使馆/领事馆预约（北京、上海、广州、成都）
        \item Termin bei deutscher Botschaft/Konsulat in China vereinbaren
        \item Submit complete application package with all documents
        \item 提交完整的申请包及所有文件
        \item Vollständiges Antragspaket mit allen Dokumenten einreichen
        \item Pay application fee (approximately €75 for national visa)
        \item 支付申请费（国家签证约75欧元）
        \item Antragsgebühr zahlen (ca. 75 Euro für Nationalvisum)
        \item Initial review typically takes 4-8 weeks
        \item 初步审查通常需要4-8周
        \item Erste Prüfung dauert typischerweise 4-8 Wochen
    \end{itemize}
    
    \item \textbf{Interview and Assessment} / \textbf{面试和评估} / \textbf{Interview und Bewertung}
    \begin{itemize}
        \item Embassy may request interview to discuss business plan and motivation
        \item 大使馆可能要求面试以讨论商业计划和动机
        \item Botschaft kann Interview zur Diskussion des Businessplans und der Motivation anfordern
        \item Prepare to explain: Why Germany? Why this business? How will you succeed?
        \item 准备解释：为什么选择德国？为什么是这个业务？您将如何成功？
        \item Vorbereitung auf Fragen: Warum Deutschland? Warum dieses Geschäft? Wie werden Sie erfolgreich sein?
        \item Local IHK or economic development agency may conduct separate assessment
        \item 当地IHK或经济发展机构可能进行单独评估
        \item Lokale IHK oder Wirtschaftsförderung kann separate Bewertung durchführen
    \end{itemize}
    
    \item \textbf{Visa Approval and Entry} / \textbf{签证批准和入境} / \textbf{Visumgenehmigung und Einreise}
    \begin{itemize}
        \item If approved, receive entry visa (valid 3-6 months, single or multiple entry)
        \item 如果获得批准，将获得入境签证（有效期3-6个月，单次或多次入境）
        \item Bei Genehmigung erhält man Einreisevisum (3-6 Monate gültig, einmalig oder mehrfach)
        \item Enter Germany within visa validity period
        \item 在签证有效期内进入德国
        \item Innerhalb der Visumsgültigkeit nach Deutschland einreisen
        \item Apply for residence permit at local Foreigners' Office (Ausländerbehörde) within 90 days
        \item 在90天内在当地外国人办公室（Ausländerbehörde）申请居留许可
        \item Innerhalb von 90 Tagen bei der örtlichen Ausländerbehörde Aufenthaltserlaubnis beantragen
        \item Initial residence permit typically valid for 1-3 years, renewable upon business success
        \item 初始居留许可通常有效期为1-3年，业务成功后可再生
        \item Erste Aufenthaltserlaubnis typischerweise 1-3 Jahre gültig, bei Geschäftserfolg verlängerbar
    \end{itemize}
\end{enumerate}

\section{Required Documents / 所需文件 / Erforderliche Dokumente}

\subsection{Personal Documents / 个人文件 / Persönliche Dokumente}
\begin{itemize}
    \item Valid passport
    \item 有效护照
    \item Gültiger Reisepass
    \item Birth certificate
    \item 出生证明
    \item Geburtsurkunde
    \item Educational certificates
    \item 教育证书
    \item Bildungszertifikate
    \item Health insurance
    \item 健康保险
    \item Krankenversicherung
\end{itemize}

\subsection{Business Documents / 商业文件 / Geschäftsdokumente}
\begin{itemize}
    \item Business plan
    \item 商业计划
    \item Businessplan
    \item Financial statements
    \item 财务报表
    \item Finanzunterlagen
    \item Proof of funds
    \item 资金证明
    \item Nachweis der Mittel
    \item Market analysis
    \item 市场分析
    \item Marktanalyse
\end{itemize}

\section{Post-Arrival Procedures / 抵达后程序 / Verfahren nach der Ankunft}

\subsection{Registration / 登记 / Anmeldung}
\begin{itemize}
    \item Residence registration (Anmeldung)
    \item 居住登记（Anmeldung）
    \item Wohnsitzanmeldung (Anmeldung)
    \item Tax registration
    \item 税务登记
    \item Steuerregistrierung
    \item Health insurance registration
    \item 健康保险登记
    \item Krankenversicherungsregistrierung
\end{itemize}

\subsection{Business Registration / 商业登记 / Geschäftsregistrierung}
\begin{itemize}
    \item Trade office registration (Gewerbeanmeldung)
    \item 贸易办公室登记（Gewerbeanmeldung）
    \item Gewerbeanmeldung
    \item Company formation
    \item 公司成立
    \item Unternehmensgründung
    \item Tax number application
    \item 税号申请
    \item Steuernummer beantragen
\end{itemize}

\section{Resources and Support / 资源和支持 / Ressourcen und Unterstützung}

\begin{itemize}
    \item German embassies and consulates in China
    \item 德国驻华大使馆和领事馆
    \item Deutsche Botschaften und Konsulate in China
    \item Immigration lawyers and consultants
    \item 移民律师和顾问
    \item Einwanderungsanwälte und -berater
    \item Government websites and portals
    \item 政府网站和门户
    \item Regierungswebsites und -portale
\end{itemize}
